---
title: "Income and Attrition"
author: "Dan Simonet"
date: "10/31/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r echo = F}
library(tidyverse)
library(sjPlot)
library(afex)
library(kableExtra)

HR_dat <- read_csv("https://raw.githubusercontent.com/DanSimonet/METRO-Workshop/main/IBM_HR_Data.csv")

HR_dat_s <- HR_dat %>% 
  filter(DistanceFromHome <= 28) %>%
   mutate(OvTenure = (YearsInCurrentRole + YearsSinceLastPromotion + YearsWithCurrManager)/3, 
          Attrition = factor(Attrition),
          Department = factor(Department),
          Gender = factor(Gender)) %>%
  select(-contains("Years")) %>%
  mutate(id = row_number())  

HR_dat_ss <- HR_dat_s %>% mutate(Department = abbreviate(Department))
inc_aov_s <- aov_car(MonthlyIncome ~ Department * Gender + Error(id),  #add Error(id)
                   data = HR_dat_ss)
```

The current report analyzes income disparities and attrition for the IBM workforce. the sample contains `r nrow(HR_dat)` employees spread across `r n_distinct(HR_dat$Department)` departments.  

### ANOVA Analysis

A 2 x 3 between-group ANOVA was conducted to evaluate income differences by department and gender. Results suggest no significant effect for gender, $F_{1, 1437} = .9038, p = .34$, or the interaction, $F_{2,1437} = .42, p = .66$, but a marginal effect for department, $F_{2, 1437} = 2.59, p = .08$, with sales (*M* = $6,953) trending towards higher monthly pay than HR (*M* = $6,655) and research (*M* = $6,289).

```{r echo = F}
inc_aov_s <- aov_car(MonthlyIncome ~ Department * Gender + Error(id),
                   data = HR_dat_ss)
kable(nice(inc_aov_s)) %>% kable_classic_2(full_width = F, position = "left")
```

### Post-Hocs 

Eyeballing narrower sub-group differences suggests male researchers make less than male salesmen ($\Delta M=\$839.36$) and female HR reps ($\Delta M=\$1,147.91$). However, Tukey's HSD test for multiple comparisons on estimated marginal means across all sub-group pairings provides further evidence of no significant pay differences.

```{r echo = F}
library(emmeans)
ph <- emmeans(inc_aov_s, c("Department", "Gender"))
plot(ph)
```

### Attrition

A logistic regression shows improvements in monthly income, work life balance, and job satisfaction all uniquely contribute to reductions in attrition. Odds ratios suggest a point increase in balance or satisfaction leads to a 22% reduction in the odds of turnover whereas making $1000 extra per month leads to a 12% drop. 

```{r echo = F}
tur_mod <- glm(Attrition ~ I(MonthlyIncome/1000) + WorkLifeBalance +  JobSatisfaction, data = HR_dat_s, family = "binomial")
tab_model(tur_mod)
```

<br></br>
The following plot shows how the probability of attrition changes across predictor levels. Those making over $15,000 are likely to stay regardless of job satisfaction and balance. However, for those making $5,000 or less a month the differences in attitudinal extremes translates into a nearly 20% reduction in probability of departure. 

```{r echo = F}
library(ggeffects)

tur_mod_1 <- glm(Attrition ~ MonthlyIncome + WorkLifeBalance +  JobSatisfaction, data = HR_dat_s, family = "binomial")

fpr <- ggpredict(tur_mod_1, c("MonthlyIncome", "WorkLifeBalance", "JobSatisfaction"))
plot(fpr)
```

